Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-07-23 manuscript/bef_stability/manuscript.Rmd

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE 
)

@@

- Number of words:
  - Abstract: 194
  - Main text: XXXX

@@

Structural Equation Modeling are able to assess multivariate relationships
between variables. It is one promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. We built the biomass
model by considering that environment
can directly affect species richness and network structure, as well as total
biomass (Figure \@ref(fig:matmet) A, \@ref(fig:sem-plot) B). 
Food web structure was shown to be constrained by species richness [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010]. 
We assumed that both species richness and network structure directly affect
total biomass (Figure \@ref(fig:matmet) A, \@ref(fig:sem-plot) B). In addition,
we assumed that species richness would indirectly affect total biomass through
network structure [@schneider_animal_2016].
We used the same model structure for temporal variability (Figure
\@ref(fig:sem-plot) A). The structural equation models used for total biomass
and its temporal stability contains respectively 4 and 6 response variables, 35
and 26 direct paths.

@@

We found that species richness increased total biomass as found in salt marshes,
grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018]. The positive effect of species
richness on total biomass might be due to resource use complementarity, which is
held by niche differentiation and interspecific facilitation
[@loreau_biodiversity_1998; @loreau_partitioning_2001; @cardinale_species_2002]
or might be due to a positive selection effect [@tilman_plant_1997;
@loreau_partitioning_2001], which leads to the dominance of the most productive
species. Previous research seems to indicate that resource complementarity might
be the main driver in fish communities [@carey_determining_2011].

Surprisingly found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006]; mesocosms (Pennekamp et al. 2018), in 
butterflies natural communities [@olivier_urbanization_2020] or even in lotic
fish communities [@franssen_annual_2011] where positive relationships were found
between temporal stability and species richness. However other studies found a
opposite relationship in natural grassland [@yang_effects_2011] or a marginal
one in natural forests [@declerck_species_2006]. A first step to identify the
mechanisms behind those contrasting patterns is be to look at the effects of
species richness on synchrony and $\overline{CV_{sp}}$, which constitute the two
components of temporal stability quantified as CV [@loreau_species_2008;
@thibaut_understanding_2013].

We found that species richness decreases synchrony, thereby increasing
stability. Species richness can reduce synchrony by decreasing of population
size, thus raising the role of demographic stochasticity on population dynamics
[@loreau_species_2008] but also in the case where species have different
responses to environmental variation [@loreau_species_2008]. While a negative
effect of species richness on synchrony is generally found in experimental
grasslands [@isbell_biodiversity_2009; @roscher_identifying_2011], no relation
has been found in natural forests [@declerck_species_2006], but negative
relationships have been found in natural communities of birds, bats, grasslands,
bats and butterflies [@bluthgen_land_2016; @olivier_urbanization_2020]. Our
study thus highlights that species richness might also decrease synchrony in
empirical trophic communities. The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
positive effect of species richness on $\overline{CV_{sp}}$ was reported in
experimental grasslands and natural bird communities [@tilman_biodiversity_2006;
@olivier_urbanization_2020], negative and neutral relationships have been found
in butterfly and bat communities. @thibaut_understanding_2013 proposed that a
strongly uneven biomass distribution across population and that the biomass
variance of population increases much faster than its mean can generate a
positive effect of species richness on population ($\overline{CV_{sp}}$)
stability. It was unexpected that the average population $\overline{CV_{sp}}$
had a stronger negative effect on biomass stability than synchrony
[@bluthgen_land_2016; @olivier_urbanization_2020], and explain why we observe a
negative effect of species on temporal stability of biomass. The lower effect of
synchrony can be explained by the the presence of generalist predator in stream
communities, which can raise synchrony between prey species which shares the
same predator [@raimondo_interspecific_2004]. Alternatively, uneveness can also
raise synchrony between species [@thibaut_understanding_2013], in line with
mechanisms which can raise positive relationships between $\overline{CV_{sp}}$
and species richness.

## Vertical diversity increased total biomass and its temporal stability

The mean trophic level was positively linked to total biomass. This result is
straightforward. As feeding interactions were partly determined by body size and
that body mass increase with body size, it is expected that the more the average
biomass is located at higher trophic level, the more the total biomass increase.
A mechanism, suggested by experimental and theoretical work, can be that a
higher trophic level is the result of a stronger top-down effect, meaning higher
predation rates and biomass fluxes to higher trophic levels
[@thebault_relationship_2006; @kratina_warming_2012], hence leading to an
increase in total biomass with trophic length. Furthermore,
@wang_biodiversity_2016 have shown that total nutrient uptake might increases
with the length of trophic chain and so results in higher total biomass.  In
contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance should translate in higher resource transfer
efficiency between prey and predators and so results in higher total biomass in
the fish compartment. A reason for such an absence of an effect might be due the
absence of information that we have about the other compartments of the
community, i.e, the primary producers, zooplankton and zoobenthos. Further
research are needed to assess the role of the trophic structure with a more
complete communities.

Mean trophic level had a positive effect on biomass stability by its negative
effect on $\overline{CV_{sp}}$, thereby increasing average species stability. It
means that species at higher trophic level are more stable than the species at
lower trophic levels. Strong top-down control can generate this pattern in pure
trophic chain [@barbier_pyramids_2018; @shanafelt_stability_2018] but it is not
clear if this theoretical prediction holds when considering omnivorous species
[@wang_biodiversity_2016] and if other mechanisms can generate the same pattern
[@thibaut_understanding_2013]. Another explanation is that the fishes at higher
trophic levels are not more stable themselves, but rather allow for stronger
top-down control and resource complementarity and thus reduce biomass
fluctuations in lower trophic levels [@rooney_structural_2006;
@neutel_reconciling_2007]. However, the ontogenic nature of the trophic networks
complicate the interpretation from the network perspective to population
fluctuations, the network structure fluctuate according to change in population
size structure. Generally, a fish will change of diet over its life span and
thus move upward in the trophic chain. The temporal dynamic of a given species
composed of several trophic species can thus be a mix of the dynamics of several
trophic levels, plus the oscillatory dynamics of the cohorts
[@townsend_population_1989; @medvinsky_modelling_2015]. Then, there is a clear
need for a better understanding of how the dynamics of pure food chain models
would be affected by inclusion of omnivory and ontogeny. We found that
connectance had no effect on temporal stability. This result previous precedent
theoretical studies on the relationship between stability and network structure
[@angelis_stability_1975; @may_will_1972; @thebault_stability_2010], although
they considered different stability concepts and metrics
[@arnoldi_resilience_2016]. Once again, it is possible that that connectance
from our networks contains too few information. Furthermore, a crucial missing
ingredient to go further is the interaction strength. Indeed, interaction
strength is a crucial ingredient of stability [@may_will_1972;
@angelis_stability_1975; @dunne_network_2006]. An alternative explanation can be
that connectance increases inhomogeneously in the food web, for instance a
higher connectance in lower trophic levels will not have the stabilising effects
conferred by higher predation and omnivory in higher trophic levels. Further
investigations on this effect is needed.

## Horizontal and vertical diversity had common, opposite and dependent effects on total biomass and its temporal stability

We found that species richness had a positive and a negative effect on
respectively total biomass and its temporal stability whereas average trophic
level had a positive effect on both. It confirms that more investigation is
needed to understand the mechanisms leading to common and opposite effects of
species richness and network structure. We found that connectance covaries
negatively with species richness [@winemiller_must_1989; @dunne_network_2006;
@thebault_stability_2010], connectance had itself no effects on both total
biomass and its temporal stability. On the other side, we found that species richness
was negatively related to mean trophic level and mean trophic level
covariated positively with total biomass and its temporal stability. Hence,
species richness had a direct positive and negative effect respectively on
total biomass and its stability but a negative and positive effect through
mean trophic level. Theoretical work have shown that species richness can be
linked to allocation of biomass toward higher trophic level
[@thebault_relationship_2006; @schneider_animal_2016; @wang_biodiversity_2016].
It indicates the importance to consider both species richness, network structure
and their dependence. It calls for further investigations on how species
diversity variation translates in network structure. Finally, common mechanisms
might be at play in mono and multi-trophic communities such as while having same or
opposite effects on total biomass and its temporal stability.

@@

We find that environmental drivers overall affect total biomass indirectly,
through changes in network structure and species richness. This contrasts with
research in plant communities that showed that total biomass was directly driven
by environmental factors [@grace_integrative_2016], and in particular by
warming, which has contrasted effects on biomass in aquatic communities
[@yvondurocher_warming_2011; @dossena_warming_2012]. Average temperature,
inverse altitude, and average stream size had indirect positive effects on total
biomass, through a positive direct effect on species richness. These effects are
well known in biogeography and in freshwater communities and are explained by a gain in species
along thermal gradients, from smaller and colder streams, which support small
food webs dominated by salmonids [@ogorman_chapter_2012;
@emmrich_geographical_2014], to wider and warmer sections of stream networks,
which contain larger food webs dominated by mesotrophic preators such as percids
[@allan_stream_2007; @emmrich_geographical_2014]. Further in line with this
pattern, we find that average stream size had a positive effect on total biomass
through a positive effect on mean trophic level, consistent with the fact that
food webs in wider streams, with a greater amount of running waters, are also
larger, and hence support more species and biomass [@allan_stream_2007]. Greater
ecosystem size might be linked to a greater diversity of resources through
habitat heterogeneity, on so higher species richness and total biomass by
surface unit, through resource complementarity. Greater diversity of resources
might also allow to maintain higher biomass in high trophic level
[@post_ecosystem_2000; @doi_resource_2009]. Overall, our results indicate that
environmental effects on total biomass operate primarily through changes in
community structure. Furthermore, the positive effect of temperature on total
biomass was of the magnitude of effects of network structure and species
richness. This adds to the mounting body of work reporting important effects of
temperature on the structure and biomass of aquatic system [e.g.
@dossena_warming_2012; @edeline_ecological_2013; @emmrich_geographical_2014;
@hattab_forecasting_2016].

@@

We aimed to assess the effect of species richness & network structure on empirical
trophic communities. While theoretical ecology has focused on the effects of
network structure on stability, empirical ecology has focused on species richness
and ignored network structure. Furthermore, both fields have largely used
different stability metrics. We showed that species richness and network
structure can have independent and sometimes opposite effects on total biomass
and its temporal stability. It opens a promising way for the understanding on
how species interactions, of different types, can affect both synchrony and $\overline{CV_{sp}}$
in different ecosystems. However, the mecanisms leading to such patterns remains
to be elucidated. The reconstruction of ecological networks based on body size
and food diet litterarure can be a first approach to document such
relationships, however it will not compensate for close evaluation of measured
interaction strength. We also believe that SEM methods are particularly helpful
to assess complex relationship between environment, community structure and
ecosystem properties in empirical data which necessitate to move from simple
bivariate relationships [@grace_integrative_2016]. We believe that the
integration of network structure and species interaction will improve our
understanding of the mechanisms driving ecosystem properties.

>>>>

knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE 
)

@@

- Number of words:
  - Abstract: 208 
  - Main text: XXXX

@@

Structural Equation Modeling are able to assess multivariate relationships
between variables. It is one promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. We built the biomass
model by considering that environment
can directly affect species richness and network structure, as well as total
biomass (Figure \@ref(fig:sem-plot) B). 
Food web structure was shown to be constrained by species richness [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010]. 
We assumed that both species richness and network structure directly affect
total biomass (Figure \@ref(fig:sem-plot) B). In addition,
we assumed that species richness would indirectly affect total biomass through
network structure [@schneider_animal_2016].
We used the same model structure for temporal variability (Figure
\@ref(fig:sem-plot) A). The structural equation models used for total biomass
and its temporal stability contains respectively 4 and 6 response variables, 35
and 26 direct paths.

@@

We found that species richness increased total biomass as found in salt marshes,
grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018]. The positive effect of species
richness on total biomass might be due to resource use complementarity, which is
held by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or might be due to a positive selection effect
[@tilman_plant_1997; @loreau_partitioning_2001], which leads to the dominance of
the most productive species. The study of @carey_determining_2011 showed that
resource complementarity might be the main driver in fish communities.

Surprisingly we found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006]; mesocosms (Pennekamp et al. 2018), in 
butterflies natural communities [@olivier_urbanization_2020] or even in lotic
fish communities [@franssen_annual_2011] where positive relationships were found
between temporal stability and species richness. However other studies found a
opposite relationship in natural grassland [@yang_effects_2011] or a marginal
one in natural forests [@declerck_species_2006]. A first step to identify the
mechanisms behind those contrasting patterns is be to look at the effects of
species richness on synchrony and $\overline{CV_{sp}}$, which constitute the two
components of temporal stability quantified as CV [@loreau_species_2008;
@thibaut_understanding_2013].

We found that species richness decreases synchrony, thereby increasing
stability. Species richness can reduce synchrony by decreasing of population
size, thus raising the role of demographic stochasticity on population dynamics
[@loreau_species_2008] but also in the case where species have different
responses to environmental variation [@loreau_species_2008]. While a negative
effect of species richness on synchrony is generally found in experimental
grasslands [@isbell_biodiversity_2009; @roscher_identifying_2011], no relation
has been found in natural forests [@declerck_species_2006], but negative
relationships have been found in natural communities of birds, bats, grasslands,
bats and butterflies [@bluthgen_land_2016; @olivier_urbanization_2020]. Our
study thus highlights that species richness might also decrease synchrony in
empirical trophic communities.

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
positive effect of species richness on $\overline{CV_{sp}}$ was reported in
experimental grasslands and natural bird communities [@tilman_biodiversity_2006;
@olivier_urbanization_2020], negative and neutral relationships have been found
in butterfly and bat communities. @thibaut_understanding_2013 proposed that a
strongly uneven biomass distribution across population and that the biomass
variance of population increases much faster than its mean can generate a
positive effect of species richness on population ($\overline{CV_{sp}}$)
stability.

It was unexpected that the average population $\overline{CV_{sp}}$
had a stronger negative effect on biomass stability than synchrony
[@bluthgen_land_2016; @olivier_urbanization_2020], and explain why we observe a
negative effect of species on temporal stability of biomass. The lower effect of
synchrony can be explained by the the presence of generalist predator in stream
communities, which can raise synchrony between prey species which shares the
same predator [@raimondo_interspecific_2004]. Alternatively, uneveness can also
raise synchrony between species [@thibaut_understanding_2013], in line with
mechanisms which can raise positive relationships between $\overline{CV_{sp}}$
and species richness.

## Mean trophic level increased total biomass and its temporal stability

The mean trophic level was positively linked to total biomass. As feeding
interactions were partly determined by body size and that body mass increase
with body size, it is expected that the more the average
biomass is located at higher trophic level, the more the total biomass increase.
A mechanism, suggested by experimental and theoretical work, can be that a
higher trophic level is the result of a stronger top-down effect, meaning higher
predation rates and biomass fluxes to higher trophic levels
[@thebault_relationship_2006; @kratina_warming_2012], hence leading to an
increase in total biomass with trophic length. Furthermore,
@wang_biodiversity_2016 have shown that total nutrient uptake might increases
with the length of trophic chain and so results in higher total biomass.

We found that species richness was negatively related to mean trophic level and 
connectance [@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010]. In
contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance should translate in higher resource transfer
efficiency between prey and predators and so results in higher total biomass in
the fish compartment. A reason for such an absence of an effect might be due the
absence of information that we have about the other compartments of the
community, i.e, the primary producers, zooplankton and zoobenthos. Further
research are needed to assess the role of the trophic structure with a more
complete communities.

Mean trophic level had a positive effect on biomass stability by its negative
effect on $\overline{CV_{sp}}$, thereby increasing average species stability. It
means that species at higher trophic level are more stable than the species at
lower trophic levels. Strong top-down control can generate this pattern in pure
trophic chain [@barbier_pyramids_2018; @shanafelt_stability_2018] but it is not
clear if this theoretical prediction holds when considering omnivorous species
[@wang_biodiversity_2016] and if other mechanisms can generate the same pattern
[@thibaut_understanding_2013]. Another explanation is that the fishes at higher
trophic levels are not more stable themselves, but rather allow for stronger
top-down control and resource complementarity and thus reduce biomass
fluctuations in lower trophic levels [@rooney_structural_2006;
@neutel_reconciling_2007].

We found that connectance had no effect on temporal
stability. This result previous precedent theoretical studies on the
relationship between stability and network structure [@angelis_stability_1975;
@may_will_1972; @thebault_stability_2010], although they considered different
stability concepts and metrics [@arnoldi_resilience_2016].  Furthermore, we did
not information about the interaction strength. Indeed, interaction strength is
a crucial ingredient of stability [@may_will_1972; @angelis_stability_1975;
@dunne_network_2006]. An alternative explanation can be that connectance
increases inhomogeneously in the food web, for instance a higher connectance in
lower trophic levels will not have the stabilising effects conferred by higher
predation and omnivory in higher trophic levels. Further investigations on this
effect is needed.

@@

We find that environmental drivers overall affect total biomass indirectly,
through changes in network structure and species richness. This contrasts with
research in plant communities that showed that total biomass was directly driven
by environmental factors [@grace_integrative_2016], and in particular by
warming, which has contrasted effects on biomass in aquatic communities
[@yvondurocher_warming_2011; @dossena_warming_2012]. Average temperature,
inverse altitude, and average stream size had indirect positive effects on total
biomass, through a positive direct effect on species richness. These effects are
well known in biogeography and in freshwater communities [@allan_stream_2007; @ogorman_chapter_2012;
@emmrich_geographical_2014].
Further in line with this pattern, we find that average stream size had a
positive effect on total biomass through a positive effect on mean trophic
level, consistent with the fact that food webs in wider streams, are also
larger, and hence support more species and biomass [@allan_stream_2007]. Greater
ecosystem size might be linked to a greater diversity of resources through
habitat heterogeneity, on so higher species richness and total biomass by
surface unit, through resource complementarity. Greater diversity of resources
might also allow to maintain higher biomass in high trophic level
[@post_ecosystem_2000; @doi_resource_2009]. Overall, our results indicate that
environmental effects on total biomass operate primarily through changes in
community structure. Furthermore, the positive effect of temperature on total
biomass was of the magnitude of effects of network structure and species
richness. This adds to the mounting body of work reporting important effects of
temperature on the structure and biomass of aquatic system [e.g.
@dossena_warming_2012; @edeline_ecological_2013; @emmrich_geographical_2014;
@hattab_forecasting_2016].

@@

We aimed to assess the effect of species richness & network structure on empirical
trophic communities. While theoretical ecology has focused on the effects of
network structure on stability, empirical ecology has focused on species richness
and ignored network structure. Furthermore, both fields have largely used
different stability metrics. We showed that species richness and network
structure can have independent and sometimes opposite effects on total biomass
and its temporal stability. Species richness had a direct positive and negative effect
respectively on total biomass and its stability but a negative and positive
effect through mean trophic level. It indicates the importance to consider both
species richness, network structure and their dependence. It calls for further
investigations on how species diversity variation translates in network
structure. Our work opens a promising way for the understanding on
how species interactions, of different types, can affect both synchrony and $\overline{CV_{sp}}$
in different ecosystems. However, the mecanisms leading to such patterns remains
to be elucidated. The reconstruction of ecological networks based on body size
and food diet litterarure can be a first approach to document such
relationships, however it will not compensate for close evaluation of measured
interaction strength. We also believe that SEM methods are particularly helpful
to assess complex relationship between environment, community structure and
ecosystem properties in empirical data which necessitate to move from simple
bivariate relationships [@grace_integrative_2016]. We believe that the
integration of network structure and species interaction will improve our
understanding of the mechanisms driving ecosystem properties.
